---
title: "Session_7_part_2"
author: "Kyle Gontjes"
date: "5/2/2022"
output: html_document
---

## R Markdown Setup Chunk
```{r setup}
require(knitr)
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
knitr::opts_chunk$set(fig.width = 15, fig.height = 10)
knitr::opts_chunk$set(comment = NA)
# Change this working directory to be your location of the session 7 folder
knitr::opts_knit$set(root.dir="~/Desktop/code_club_curriculum/Session_7/") 
```

## Agenda
0. Set up the R Environment
1. Load data
2. Perform Initial Visualization of the Phylogeny Using the ggtree Package
3. Load in Previous Session's Final ST258 Tree, Data, and Visualization
4. Try Different Phylogeny Layouts
5. Re-do Session 7's Phylogeny as Circular Tree Using Theme Objects

## Part 0: Set up the R Environment   
Note: This code chunk loads the required packages for this code club session. Instead of using the library function consecutively for each package, I often use this lapply function strategy to load my packages at the start of my R markdown file.
```{r}
# Packages used 
  # Cran packages
  cran_packages <- c("BiocManager","phytools","ggplot2","tidyverse","readxl","cowplot","printr","randomcoloR","magick")
  # BiocManager packages
  BiocManager_packages <- c("ggtree","ggnewscale")
  
# Functions to install packages (if not installed already)
  # lapply(cran_packages,install.packages,character.only=T)
  # lapply(BiocManager_packages,BiocManager::install,character.only=T)

# Functions to load packages
lapply(cran_packages,library,character.only=T)
lapply(BiocManager_packages,library,character.only=T)
```

## Part 1: Load the .treefile Into the R Environment
```{r}
# Note: If you choose to work in the console and not knit your .Rmd, set your working directory to YOUR location of the session 7 folder
setwd("~/Documents/code_club_curriculum/Session_7/")
# Load .treefile using the ape package ()
tree <- read.tree(file="./2021_02_12_08_34_28_KPNIH1_genome_aln_w_alt_allele_unmapped.treefile")
```

## Part 2: Perform Initial Visualization of the Tree Using ggtree
#### See the ggtree full vignette at "https://yulab-smu.top/treedata-book/"
```{r}
# Visualize tree object using the ggtree function
  # Vignette
  vignette("ggtree")
  #ggtree function help page
  ??ggtree
  # Visualize tree using ggtree
  ggtree(tree)
```

## Part 3: Load in Previous Session's Final ST258 Tree, Data, and Visualization
```{r}
# Load ST258 Datset
CRKP_258 <- readRDS("./CRKP_258_sorted.RDS")
#Load ST258 tree
tree_258 <- read.tree("./st_258.treefile")
  # Visualize tree using ggtree
  ggtree(tree_258)
#Load Initial Visualization
initial_visualizaiton <- readRDS("./final_session_7_phylogeny.RDS")
  #View initial visualization
  initial_visualizaiton
  
#Question: Can Anyone think of ways to improve the figure?
```

## Part 4: Try Different Phylogeny Layouts
```{r}
# Review: Using Different Tree Layouts
  # Default (layout = "rectangular")
  ggtree(tree_258)
  # Slanted
  ggtree(tree_258,layout = "slanted")
  # Dendogran
  ggtree(tree_258,layout = "dendrogram")
  # Ape
  ggtree(tree_258,layout = "ape")
  # No Branch Length
  ggtree(tree_258,layout="rectangular",branch.length = 'none')

# Circular Phylogenies
ggtree(tree_258,layout = "circular")
  #Add Scale to Tree
  ggtree(tree_258,layout = "circular") + geom_treescale()
  #Ladderize
  ggtree(tree_258,layout = "circular",ladderize=T)  + geom_treescale()
```

## Part 5: Re-do Session 7's Phylogeny as Circular Tree with Environments
# Task 1: Add Source Using gheatmap
# Task 2: Create Consistent Theme
# Task 3: Add Colistin Resistance
```{r}
# Previous Tree
initial_visualizaiton

#Re-Introducing ggtree scales
  help(ggnewscale)
  help(scale_fill_manual)

#Source Scale
  #Values
  table(CRKP_258$Source)
  #Colors
  source_colors <- distinctColorPalette(4)
    # View Color Palette Chosen Using Pie Function
    source_colors
    pie(rep(1,length(source_colors)),col=source_colors)
  #Create Named Vector
  palette_source = c("Blood" = source_colors[[1]],"Respiratory" = source_colors[[2]],"Urine" = source_colors[[3]],"Wound"=source_colors[[4]])
  palette_source

#Initial Phylogeny
p0 <- ggtree(tree_258,layout="circular")

#P1 - Source Gheatmap
p1 <-  gheatmap(p0,CRKP_258 %>% select(Source),colnames_position="top", colnames_angle=45, colnames_offset_y=0.25, width=.15,hjust=0,font.size=5,color=NA) + scale_fill_manual(values=palette_source,name="Source") + ylim(NA,500) 
  #Use Theme to Put Legend at Bottom
  p1 + theme(legend.position="bottom")
 
#Option #2: Create Legend Object
source_legend <- scale_fill_manual(breaks=c("Blood","Respiratory","Urine","Wound"),values = c(2,3,4,6),labels=c("Blood","Resp","Urine","Wound"),name = "Source",guide = guide_legend(title.position = "top", label.position = "right",ncol=2,order=1))

p1.1 <- gheatmap(p0,CRKP_258 %>% select(Source),colnames_position="top", colnames_angle=45, colnames_offset_y=0.25, width=.15,hjust=0,font.size=5,color=NA) + source_legend + theme(legend.position="bottom") + ylim(NA,500) 

#Task #2: Create Consistent Theme
consistent_theme <- theme(legend.position = 'bottom',legend.direction="horizontal", legend.justification = "center", legend.key = element_rect(colour = c('black')))
p1
  #Add Consistent Theme + Source Legend to Phylogeny
  p1.2 <- p1 + source_legend + consistent_theme
  p1.2

#Task 3: Add Colistin Resistance
#Use new_scale_fill to add new scale for colistin data
p2 <- p1.2 + new_scale_fill()
p3 <- gheatmap(p2, CRKP_258 %>% select(Colistin), colnames_position="top",  colnames_angle=45, colnames_offset_y=0.25, width=.15, hjust=0, font.size=5, offset=.0000055)  + consistent_theme + ylim(NA, 500)
p3

#Make Black & White Scale
  #Values
  table(CRKP_258$Colistin)
  #scale
  colistin_scale <- scale_fill_manual(breaks = c("Non-Susceptible","Susceptible"),values=c("black","white"),labels = c("Non-Susceptible","Susceptible"),name="Resistance Profile",guide = guide_legend(title.position = "top", label.position = "right",nrow=2,order=2))

#Final Tree with colistin scale
final_phylogeny <- p3 + colistin_scale + consistent_theme + geom_treescale(x=.00001,y=-25,offset=1,fontsize=3)

#Trim Circular Plot Using Magickr
f <- tempfile(fileext=".png")
ggsave(filename = f, plot = final_phylogeny, width=7, height=7)

x <- image_read(f, density=300)
y <- image_trim(x)
y
```