---
title: "Session_7_part_2"
author: "Kyle Gontjes"
date: "5/2/2022"
output: html_document
---

## R Markdown Setup Chunk
```{r setup}
require(knitr)
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
knitr::opts_chunk$set(fig.width = 15, fig.height = 10)
knitr::opts_chunk$set(comment = NA)
# Change this working directory to be your location of the session 7 folder
knitr::opts_knit$set(root.dir="~/Desktop/code_club_curriculum/Session_7/") 
```

## Agenda
0. Set up the R Environment
1. Load data
2. Perform Initial Visualization of the Phylogeny Using the ggtree Package
3. Load in Previous Session's Final ST258 Tree, Data, and Visualization
4. Try Different Phylogeny Layouts
5. Re-do Session 7's Phylogeny as Circular Tree Using Theme Objects

## Part 0: Set up the R Environment   
Note: This code chunk loads the required packages for this code club session. Instead of using the library function consecutively for each package, I often use this lapply function strategy to load my packages at the start of my R markdown file.
```{r}
# Packages used 
  # Cran packages
  cran_packages <- c("BiocManager","phytools","ggplot2","tidyverse","readxl","cowplot","printr","randomcoloR","magick")
  # BiocManager packages
  BiocManager_packages <- c("ggtree","ggnewscale")
  
# Functions to install packages (if not installed already)
  # lapply(cran_packages,install.packages,character.only=T)
  # lapply(BiocManager_packages,BiocManager::install,character.only=T)

# Functions to load packages
lapply(cran_packages,library,character.only=T)
lapply(BiocManager_packages,library,character.only=T)
```

## Part 1: Load the .treefile Into the R Environment
```{r}
# Note: If you choose to work in the console and not knit your .Rmd, set your working directory to YOUR location of the session 7 folder
setwd("~/Documents/code_club_curriculum/Session_7/")
# Load .treefile using the ape package
```

## Part 2: Perform Initial Visualization of the Tree Using ggtree
#### See the ggtree full vignette at "https://yulab-smu.top/treedata-book/"
```{r}
# Visualize tree object using the ggtree function
  #Vignette
  vignette("ggtree")
  #ggtree function help page
  ??ggtree
  #Visualize tree using ggtree
  
```

## Part 3: Load in Previous Session's Final ST258 Tree, Data, and Visualization
```{r}
#Load ST258 Datset

#Load ST258 tree

  #Visualize tree using ggtree

#Load Initial Visualization

  #View initial visualization

  
#Question: Can Anyone think of ways to improve the figure?
```

## Part 4: Try Different Phylogeny Layouts
```{r}
# Review: Using Different Tree Layouts
  # Default (layout = "rectangular")

  # Slanted
 
  # Dendogran
 
  # Ape 

  # No Branch Length

# Circular Phylogeny

  #Add Scale to Tree

  #Ladderize

```

## Part 5: Re-do Session 7's Phylogeny as Circular Tree with Environments
# Task 1: Add Source Using gheatmap
# Task 2: Create Consistent Theme
# Task 3: Add Colistin Resistance
```{r}
# Previous Tree


#Re-Introducing ggtree scales
  help(ggnewscale)
  help(scale_fill_manual)

#Source Scale
  #Values
  table()
  #Colors
  
    # View Color Palette Chosen Using Pie Function
  
  #Create Named Vector

#Initial Phylogeny

#P1 - Source Gheatmap
  
  #Use Theme to Put Legend at Bottom
 
###Option #2: Create Legend Object

##Task #2: Create Consistent Theme
  
  #Add Consistent Theme + Source Legend to Phylogeny

#Task 3: Add Colistin Resistance
#Use new_scale_fill to add new scale for colistin data

#Make Black & White Scale
  #Values
  table()
  #scale

#Final Tree with colistin scale

###TIP: Trim Circular Plot Using Magickr
```