---
title: "Code Club Session 7"
author: "Kyle Gontjes"
date: "10/29/2021"
output: html_document
---

## R Markdown Setup Chunk
```{r setup}
require(knitr)
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
knitr::opts_chunk$set(fig.width = 15, fig.height = 10)
knitr::opts_chunk$set(comment = NA)
# Change this working directory to be your location of the session 7 folder
knitr::opts_knit$set(root.dir="~/Desktop/code_club_curriculum/Session_7/") 
```

## Agenda
0. Set up the R Environment
1. Load the .treefile Into the R Environment
2. Perform Initial Visualization of the Phylogeny Using the ggtree Package
3. Clean the Phylogeny by Removing Tips and Midpoint Rooting
4. Explore Whether Subseting Tree Based on Sequence Type of Interest is Valuable
5. Subset Tree Based on Sequence Type
6. Visualize Colistin Resistance Using Gheatmap
7. Using ggnewscale to plot heatmap with variables that have different data structures

## Part 0: Set up the R Environment   
Note: This code chunk loads the required packages for this code club session. Instead of using the library function consecutively for each package, I often use this lapply function strategy to load my packages at the start of my R markdown file.
```{r}
# Packages used 
  # Cran packages
  cran_packages <- c("BiocManager","phytools","ggplot2","tidyverse","readxl","randomcoloR","cowplot","printr")
  # BiocManager packages
  BiocManager_packages <- c("ggtree","ggnewscale")
  
# Functions to install packages (if not installed already)
  # lapply(cran_packages,install.packages,character.only=T)
  # lapply(BiocManager_packages,BiocManager::install,character.only=T)

# Functions to load packages
lapply(cran_packages,library,character.only=T)
lapply(BiocManager_packages,library,character.only=T)
```

## Part 1: Load the .treefile Into the R Environment
```{r}
# Note: If you choose to work in the console and not knit your .Rmd, set your working directory to YOUR location of the session 7 folder
setwd("~/Desktop/code_club_curriculum/Session_7/")
# Load .treefile using the ape package ()
tree <- read.tree(file="./2021_02_12_08_34_28_KPNIH1_genome_aln_w_alt_allele_unmapped.treefile")

# Question: What type of object is the .treefile?
# Answer: A list of the class "phylo." 
class(tree)
```

## Part 2: Perform Initial Visualization of the Tree Using ggtree
#### See the ggtree full vignette at "https://yulab-smu.top/treedata-book/"
```{r}
# Visualize tree object using the ggtree function
  # Help page documentation for ggtree function
  help(ggtree)
  # Use the ggtree function to view phylogeny
  ggtree(tree)

# Exercise: Add a (1) treescale using geom_treescale and (2) overlay a title using a ggplot2 functions 
ggtree(tree) + geom_treescale(x=0,y=-10,offset=1) + ggtitle("Initial Phylogeny")
  
# Question: After adding the scale to the phylogeny, what are your takeaways? Does this tell you anything about the genomic diversity of the isolates used in the tree? 

# Answer: Notice the differential branch lengths in the phylogeny, specifically that there are ~ 400 isolates with a smaller distance between edges. 
```

## Part 3: Clean the Phylogeny by (1) Removing Tips and (2) Performing Midpoint Rooting
```{r}
# Exercise: Identify the location of the tip labels for this tree
head(tree$tip.label)

# Drop Reference group using the treeio package's drop.tip function
tree_wo_root <- treeio::drop.tip(tree,'gi_661922017_gb_CP008827.1_')

# View tree without reference group
ggtree(tree_wo_root)  + ggtitle("Tree Without the Reference Group")

# Question: Would midpoint rooting be valuable for this phylogeny? 
  #Answer: Midpoint root a phylogeny involves locating the longest path between any two tips and putting the root at that location
# Midpoint root using phytools rooting tool
tree_final <- phytools::midpoint.root(tree_wo_root)

# View midpoint rooted tree again
ggtree(tree_final)  + ggtitle("Midpoint Rooted Tree")

# Question: What is the difference in the tree_wo_root and tree_final? 
  # Plot the ggtree images side-by-side using the cowplot package's plot_grid function
  plot_grid(ggtree(tree_wo_root), ggtree(tree_final), ncol=2,labels = c('Without Midpoint Rooting','Midpoint Root'))
```

## Part 4: Explore Whether Subseting the Tree on Sequence Type of Interest is Valuable
```{r}
# Import CRKP Metadata
CRKP_metadata <- as.data.frame(read_xlsx('./CRKP_metadata.xlsx'))

# Exercise: Characterize the dataset
  #1. What are the variables in this dataframe?
    # Answer: isolate_no, LTACH, Region, Source, ST (Sequence Type), and Colistin (Susceptibility)
  names(CRKP_metadata)
  #2. How many isolates are in this dataframe?
    # Hint: Each row = a unique isolate
    # Answer: 458 isolates
  nrow(CRKP_metadata)
  #3. Does the number of isolates in the metadata dataframe equal the number of isolates in the phylogeny?
    # Answer: No. Not all isolates were sequenced and included in the phylogeny. 
  nrow(CRKP_metadata) == tree$Nnode
  #4. Does the row names of the dataframe match the tip labels in the tree? 
    # Answer: No. We must ensure that row names of the dataframe match the tip labels of the phylogeny to ensure proper mapping of the data to the phylogeny. 
  head(rownames(CRKP_metadata))
  head(tree_final$tip.label)

# Create rownames for sorting the dataframe to match the tiplabels of the tree
rownames(CRKP_metadata) <- paste0("PCMP_H",CRKP_metadata$isolate_no)
  
# Subset and sort the dataframe using the match function
CRKP_metadata_final <- as.data.frame(CRKP_metadata[match(as.vector(tree_final$tip.label), row.names(CRKP_metadata)), ])

# Question: What is the distribution of sequence types within the dataset?
  # Data table:
table(CRKP_metadata_final$ST)
  # Bar plot  
ggplot(data = CRKP_metadata_final, aes(x = ST)) +
    geom_bar()

# Add tips to the tree to visualize distribution of sequence types across the phylogeny
  # Create tip for sequence type
  ST_tip <- c(CRKP_metadata_final$ST,rep(NA,tree_final$Nnode))
  # Visualize tree with sequence type as tips 
  ggtree(tree_final) +  geom_tippoint(aes(color=ST_tip,legend_title="Sequence Types"))
  # Question: How is sequence type distributed across the phylogeny?
    # Answer: The clade with a relatively lower distances between tips are predominantly ST258 isolates
```

## Part 5: Subset Tree Based on Sequence Type
```{r}
# Subset the tree to include only ST258 isolates
  # Create ST258 dataframe
  CRKP_258 <- subset(CRKP_metadata_final,ST == "258")
  # Subset phylogeny to include only ST258 isolates
  tree_258 <- drop.tip(tree_final,tree_final$tip.label[!tree_final$tip.label %in% rownames(CRKP_258)])
  # Sort dataframe by tip label
  CRKP_258_sorted <- CRKP_258[match(tree_258$tip.label, rownames(CRKP_258)), ] 
  
# Visualize ST258 phylogeny
ggtree(tree_258)

# Visualize ST258 phylogeny using different layouts
  # Default (layout = "rectangular")
  ggtree(tree_258)
  # Circular
  ggtree(tree_258,layout = "circular")
  # Slanted
  ggtree(tree_258,layout = "slanted")
  # Dendogran
  ggtree(tree_258,layout = "dendrogram")
  # Ape
  ggtree(tree_258,layout = "ape")
  # No Branch Length
  ggtree(tree_258,layout="rectangular",branch.length = 'none')

# Exercise 1: Create a tip label for Source
  # 1. Tabulate source data
  table(CRKP_258_sorted$Source)
  # 2. Create tip for Source
  Source_tip <- c(CRKP_258_sorted$Source,rep(NA,tree_258$Nnode))
  # 3. Visualize
  ggtree(tree_258) +  geom_tippoint(aes(color=Source_tip),size=2) + labs(color="Source")
  # 4. Store as tree_258_source
  tree_258_source <- ggtree(tree_258) +  geom_tippoint(aes(color=Source_tip),size=2) + labs(color="Source")
  
# Exercise 2: Create a tip label for Colistin Resistance
  # 1. How many isolates are susceptible to colistin? 
    # Answer: 275 (66.9%)
  table(CRKP_258_sorted$Colistin)
  # 2. Create tip for colistin resistance
  Colistin <- c(CRKP_258_sorted$Colistin,rep(NA,tree_258$Nnode))
  # 3. Visualize and improve size
  ggtree(tree_258) +  geom_tippoint(aes(color=Colistin,legend_title="Colistin Resistance"),size=2)

# Exercise 3: Create a phylogeny with geom_tippoint (1) color = source and (2) shape = colistin resistance
  ggtree(tree_258) +  geom_tippoint(aes(color=Source_tip,shape=Colistin,legend_title="Region"),size=2)+ labs(color="Source")
```

## Part 6: Visualize colistin Resistance Using gheatmap
```{r}
# gheatmap Help Page
help(gheatmap)

# Initial gheatmap of Colistin Resistance (without formatting)
gheatmap(ggtree(tree_258),CRKP_258_sorted %>% select(Colistin))+ ylim(NA, 450) 

# Stepwise formatting of gheatmap
  # Change Heatmap Structure: Angle, Offset, Width, Justification, and Font Size
  gheatmap(ggtree(tree_258), CRKP_258_sorted %>% select(Colistin), colnames_position="top", colnames_angle=90, colnames_offset_y=0.25, width=.05,hjust=0,font.size=5) + ylim(NA, 450)
  # Change Heatmap Fill Color and Legend Name: scale_fill_manual
  gheatmap(ggtree(tree_258), CRKP_258_sorted %>% select(Colistin), colnames_position="top", colnames_angle=90, colnames_offset_y=0.25, width=.05,hjust=0,font.size=5) + scale_fill_manual(values=c("black","white"),name="Resistance Profile") + ylim(NA, 450)
  # Change Legend Information: Theme 
  gheatmap(ggtree(tree_258), CRKP_258_sorted %>% select(Colistin), colnames_position="top", colnames_angle=90, colnames_offset_y=0.25, width=.05,hjust=0,font.size=5) + scale_fill_manual(values=c("black","white"),name="Resistance Profile") + theme(legend.position ='bottom',legend.key = element_rect(colour=c('black')), legend.direction="horizontal",legend.title = element_text(size=14),legend.text = element_text(size=8)) + ylim(NA, 475)  

# Exersise: Add stored ST258 phylogeny that has source tips to the gheatmap function created above
  # Hint: Format of gheatmap function = gheatmap(phylogeny, data, options)
gheatmap(tree_258_source, CRKP_258_sorted %>% select(Colistin), colnames_position="top", colnames_angle=90, colnames_offset_y=0.25, width=.05,hjust=0,font.size=5) + scale_fill_manual(values=c("black","white"),name="Resistance Profile") + theme(legend.position ='bottom',legend.key = element_rect(colour=c('black')), legend.direction="horizontal",legend.title = element_text(size=14),legend.text = element_text(size=8)) + ylim(NA,475)   
```

## Part 7: Using ggnewscale to plot heatmap with variables that have different data structures
```{r}
# Exercise: Plot source and resistance (Colistin) in the same gheatmap
gheatmap(ggtree(tree_258),CRKP_258_sorted %>% select(Source,Colistin))+ ylim(NA, 450)
  # Question: What is 'wrong' about the figure above? How can you fix it? 
  # Answer: Note the issue with the legend. Each variable has different values. Thus, they should have different scales!

#Introducing: ggnewscale's new_scale_fill function!
help(new_scale_fill)

# Generate of Unique Colors
  # Generate Distinct Color Palette Using the distinctColorPallete() function
    # Note: You can also use the randomColor function to generate a vector of random colors. 
  palette_source <- distinctColorPalette(4)
  # View Color Palette Chosen
  pie(rep(1,length(palette_source)),col=palette_source)

# Generate Heatmap Using ggnewscale
  # Phylogeny with Source Heatmap Column
  p1 <- gheatmap(ggtree(tree_258),CRKP_258_sorted %>% select(Source),colnames_position="top", colnames_angle=90, colnames_offset_y=0.25, width=.05,hjust=0,font.size=6) + scale_fill_manual(values=palette_source,name="Source") + ylim(NA,500)
  print(p1)
  #Use new_scale_fill to add optionc 
  p2 <- p1 + new_scale_fill()
  #Add Resistance Column to Heatmap
  p3 <- gheatmap(p2, CRKP_258_sorted %>% select(Colistin), colnames_position="top",  colnames_angle=90, colnames_offset_y=0.25, width=.05, hjust=0, font.size=6, offset=.0000018)  + theme(legend.position ='bottom',legend.key = element_rect(colour=c('black')), legend.direction="horizontal",legend.title = element_text(size=14),legend.text = element_text(size=8))+ geom_treescale(x=0,y=-15,offset=1)   + scale_fill_manual(values=c("black","white"),name="Resistance Profile") + ylim(NA, 500)
  print(p3)
```